---
title: "Introduction to o2mod.transphylo"
author: "Xavier Didelot"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width=8, 
  fig.height=5, 
  fig.path="figs-introduction/"
)
```

##Initialisation

```{r}
rm(list=ls())
library(ape)
library(outbreaker2)
library(TransPhylo)
library(o2mod.transphylo)
set.seed(0)
```

##Data

We consider a random timed tree with 5 leaves. Time is measured discretely, in units of days.

```{r}
nsam <- 5
phy <- rtree(nsam, br=rexp)
phy$tip.label <- 1:nsam
phy$edge.length <- round(phy$edge.length * 365)
plot(phy)
axisPhylo()
```

The outbreaker2 dataset includes the sampling dates and the generation time
distribution.  The genetic data is empty, but we attach the phylogenetic tree
itself.  This is because the TransPhylo model works by inferring the
transmission tree based on a phylogenetic tree rather the genetic data
itself. For real datasets, the phylogenetic tree needs to be reconstructed, for
example using BEAST.

```{r}
library(distcrete)
dates <- dist.nodes(phy)[nsam+1,1:nsam]
si <- distcrete("gamma", shape = 2, scale = 0.5 * 365, interval = 1L, w = 0)
w <- si$d(1:(2*max(dates)))
data <- outbreaker_data(dates = dates, w_dens = w, dna = as.DNAbin(matrix('A',nsam,1)))
data$ptree <- ptreeFromPhylo(phy,max(dates))
```

##Results

```{r}
res=outbreaker_transphylo(data)
```

```{r}
plot(res,type='alpha',burnin=0.1*length(res$step))
```

Infection dates:
```{r}
plot(res, type = 't_inf', burnin = 0.1 * max(res$step))
```

