---
title: "Introduction to o2mod.transphylo"
author: "Xavier Didelot"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width=8, 
  fig.height=5,
  fig.path = 'figs-introduction/'
)
```

## Initialisation

```{r}
library(ape)
library(outbreaker2)
library(TransPhylo)
library(o2mod.transphylo)
set.seed(0)
```

## Data

We consider a random dated tree with five leaves. Time is measured discretely, in units of days.

```{r}
nsam <- 5
phy <- rtree(nsam, br = rexp)
phy$tip.label <- 1:nsam
phy$edge.length <- round(phy$edge.length * 365)
plot(phy)
axisPhylo()
```

The outbreaker2 dataset includes the sampling dates and the generation time
distribution.  The genetic data is empty, but we attach the phylogenetic tree
itself.  This is because the TransPhylo model works by inferring the
transmission tree based on a phylogenetic tree rather the genetic data
itself. For real datasets, the phylogenetic tree needs to be reconstructed, for
example using BEAST.

```{r}
library(distcrete)
dates <- dist.nodes(phy)[nsam+1,1:nsam]
si <- distcrete("gamma", shape = 2, scale = 0.5 * 365, interval = 1L, w = 0)
w <- si$d(1:(2*max(dates)))
data <- outbreaker_data(dates = dates, w_dens = w, dna = as.DNAbin(matrix('A',nsam,1)))
data$ptree <- ptreeFromPhylo(phy, max(dates))
```

## Results

Let's run outbreaker2 using the o2mod.transphylo module:
```{r}
res <- o2mod.transphylo(data)
```
Trace of the posterior probability:
```{r}
plot(res)
```

Ancestry matrix:
```{r}
plot(res,type = 'alpha', burnin = 0.1*length(res$step))
```

Infection dates:
```{r}
plot(res, type = 't_inf', burnin = 0.1 * max(res$step))
```

We can extract the transmission tree in the last state of the MCMC, combine it with the phylogeny and visualize the resulting colored tree:

```{r}
tinf <- rep(0, nsam)
alpha <- rep(0, nsam)
for (i in 1:nsam) {
  alpha[i] <- tail(res[[8 + i]],1)
  tinf[i] <- tail(res[[8 + nsam+i]],1)
}

alpha[which(is.na(alpha))] <- 0
ttree <- list(ttree = cbind(tinf, data$dates, alpha),
              nam = data$ptree$nam)
ctree <- combine(ttree, data$ptree)
plotCTree(ctree)
```
